Generate data for multispecies modeling, including diet!

### 2020-09-11 NOBAmsdiets.Rmd
load function finally works

Links consulted: 
[awk is the right thing](https://www.tim-dennis.com/data/tech/2016/08/09/using-awk-filter-rows.html)
[awk to delete lines summing to 0](https://unix.stackexchange.com/questions/170588/delete-lines-that-sum-to-zero)  
[awk a compressed file leaving it compressed](https://unix.stackexchange.com/questions/498587/how-to-extract-information-using-awk-on-gz-files-without-storing-the-uncompress)

@@

```
zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 > DetDietHead.txt
zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 | gzip > DetDietHead.gz
cat DetDietHead.gz DetDiet.gz > NOBADetDiet.gz
```

@@

#keeps failing with vector memory exhausted (limit reached?)
#gets to the step of eliminating 0 rows
#try awk in a system command
# this test works but R overhead isnt worth it, faster run from terminal
#system("zcat < atlantisoutput/NOBA_march_2020/nordic_runresults_01DetailedDietCheck.txt.gz | awk '{print NF}' | uniq")

diet <- data.table::fread(file.path(here("atlantisoutput",
                                                  "NOBA_march_2020"), 
                                                "NOBADetDiet.gz"))

@@

  # Change column order--NOT DATA TABLE SAFE
  #diet <- diet[, c("Predator", "Cohort", "Time", "Box", "Layer",
  #  names(diet)[which(!names(diet) %in% c("Predator", "Cohort", "Time", "Box", "Layer"))])]

  data.table::setcolorder(diet, c("Predator", "Cohort", "Time", "Box", "Layer"))

  #should do the same thing faster
  #diet <- diet[as.logical(abs(as.matrix(diet[,-c(1:5)])) %*% rep(1L,ncol(diet[,-c(1:5)]))), ]

  # dataframe
  #diet <- as.data.frame(diet)

@@

@@

# function not working so test piecewise again
# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
#                                                  "NOBA_march_2020"), 
#                            "NOBADetDiet.gz", 
#                            fgs = fgs)

# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
#                                                  "NOBA_march_2020"), 
#                            "nordic_runresults_01DetailedDietCheck.txt.gz", 
#                            fgs = fgs)

# detailedpartdt <- data.table::fread(file.path(here("atlantisoutput",
#                                                  "NOBA_march_2020"), 
#                            "nordic_runresults_01DetailedDietCheck.txt.gz"),
#                            nrows = 20000
#                            )

@@

Still having problems with the function after editing

@@

# still not working but works line by line
detdiet <- load_detailed_diet_comp(here("atlantisoutput","NOBA_march_2020"), 
                                   "NOBADetDiet.gz", 
                                   fgs = fgs)

@@

>>>>

Links consulted: 
[awk is the right thing](https://www.tim-dennis.com/data/tech/2016/08/09/using-awk-filter-rows.html)

[awk to delete lines summing to 0](https://unix.stackexchange.com/questions/170588/delete-lines-that-sum-to-zero)

[awk a compressed file leaving it compressed](https://unix.stackexchange.com/questions/498587/how-to-extract-information-using-awk-on-gz-files-without-storing-the-uncompress)

@@

```
zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 > DetDietHead.txt
zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 | gzip > DetDietHead.gz
cat DetDietHead.gz DetDiet.gz > NOBADetDiet.gz
```
Now we diagnose the load function again and try to optimize with data.table functions.

@@

#keeps failing with vector memory exhausted (limit reached?)
#gets to the step of eliminating 0 rows
#try awk in a system command
# this test works but R overhead isnt worth it, faster run from terminal
#system("zcat < atlantisoutput/NOBA_march_2020/nordic_runresults_01DetailedDietCheck.txt.gz | awk '{print NF}' | uniq")
# see notes above for preprocessing step, here debug new load_detailed_diet function

dir <- here("atlantisoutput","NOBA_march_2020") 
file_diet <-  "NOBADetDiet.gz"

  if (is.null(dir)) {
    diet.file <- file_diet
  } else {
    diet.file <- file.path(dir, file_diet)
  }
  diet <- data.table::fread(diet.file)

@@

  # Change column order NOT DATA TABLE SAFE
  #diet <- diet[, c("Predator", "Cohort", "Time", "Box", "Layer",
  #  names(diet)[which(!names(diet) %in% c("Predator", "Cohort", "Time", "Box", "Layer"))])]

  # Change column order
  data.table::setcolorder(diet, c("Predator", "Cohort", "Time", "Box", "Layer"))

@@

  # # Convert to tidy dataframe to allow joining/merging with other dataframes.
  # diet <- tidyr::gather_(data = diet, key_col = "prey", value_col = "dietcomp",
  #   #gather_cols = names(diet)[(which(names(diet) == "Updated") + 1):NCOL(diet)])
  #   gather_cols = names(diet)[(which(names(diet) %in% fgs$Code))])

@@

@@

Using `data.table` improves speed for the `atantisom::load_detailed_diet_comp()` function. But it turns out you need to import data.table within a package in *both* the DESCRIPTION and NAMESPACE files [instructions!](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-importing.html) and now this beast of a function works.

It takes minutes to load the detailed diet, even with pre-processing with awk as above, so saving the output is a good idea. We save only for our 10 species to streamline things a bit.

@@

# still not working but works line by line--fixed
# N.B. data.table wants to be in both imports and namespace within a function!
detdiet <- load_detailed_diet_comp(here("atlantisoutput","NOBA_march_2020"), 
                                   "NOBADetDiet.gz", 
                                   fgs = fgs)

@@

saveRDS(ms_detaileddiet, file.path(d.name, paste0(scenario.name, "ms_detaileddiet.rds")))

All this was just to see what the regional diet data looks like compared with global diet comps from DietCheck.txt.

```{r detdiet-vis}

# look at a single species/age combo from above

# how many polygons included? how many layers per polygon have data?

# are there differences by layer within a polygon?

# are there differences by polygon? (aggregate layers)
